---
title: "Clean up BPD job codes + titles"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(rvest)
```

I'm trying to reconcile job titles [scraped from the HR site](./rselenium_job_codes.R) with the titles in the public employees rosters. Main issue was just abbreviations that didn't match, e.g. "supv" or "supervi" instead of "supervisor." HR site had full words, not abbreviations, so I cleaned roster titles to match HR.

```{r}
replacements <- c(
  `\\bsupv\\b` = "supervisor",
  `\\basst\\b` = "assistant",
  `\\bprgm\\b` = "program",
  `\\bprog\\b` = "program",
  `\\bmech\\b` = "mechanic",
  `a & p` = "air & powerplant",
  `a&$` = "air & powerplant",
  `-air&power` = " air & powerplant",
  `\\bserv\\b` = "services",
  `\\bsrvc\\b` = "services",
  `\\bspec\\.?\\b` = "specialist",
  `\\bphoto\\b` = "photographer",
  `\\btech\\b` = "technician",
  `\\bsupervi\\b` = "supervisor",
  `\\boff\\b` = "officer",
  `\\bmaint\\b` = "maintenance",
  `\\basso\\b` = "associate",
  `\\bsr\\b` = "senior",
  `\\bassoc\\b` = "associate",
  `\\bair &` = "airframe &"
  )
```



```{r}
hr_codes <- read_csv("hr_scraped_job_codes.csv") %>%
  mutate(title_clean = title %>%
           str_to_lower() %>%
           str_replace_all("-", " ") %>%
           str_replace("filght", "flight") %>%
           str_replace_all("\\s{2,}", " ")) %>%
  rename(hr_code = code) %>%
  select(hr_code, title_clean)

nrow(hr_codes)
```

Cleaning up titles from rosters and taking distinct combinations of messy title, clean title, and code--gets down to 165 unique combos.

```{r}
rosters <- file.path("..", "rosters") %>%
  list.files(path = ., pattern = "*.xlsx", full.names = T) %>%
  set_names() %>%
  set_names(str_extract, "[\\w\\d]+(?=\\.xlsx$)") %>%
  map(readxl::read_excel) %>%
  map(janitor::clean_names) %>%
  map(select, emplid, job_code, job_title, starts_with("seq")) %>%
  map(rename_all, str_remove, "(?<=seq).+$")


rost_codes <- rosters %>%
  bind_rows(.id = "file") %>%
  mutate(title_clean = job_title %>%
           str_to_lower() %>%
           str_replace_all(replacements) %>%
           str_remove_all("\\.") %>%
           str_replace_all("-", " ")) %>%
  arrange(title_clean) %>%
  distinct(job_code, title_clean, job_title)

head(rost_codes)
nrow(rost_codes)
```

Now there's only 3 job codes where clean versions of titles from roster & HR don't match, and they seem to be from compound jobs.

```{r}
rost_codes %>%
  left_join(hr_codes, by = c("job_code" = "hr_code"), suffix = c("_from_roster", "_from_hr")) %>%
  select(job_code, title_clean_from_roster, title_clean_from_hr) %>%
  filter(title_clean_from_roster != title_clean_from_hr) 
```

Figure out which job titles are sworn officers, based on [Wikipedia](https://en.wikipedia.org/wiki/Baltimore_Police_Department#Rank_structure_and_insignia). The only one that seems off is that Wikipedia doesn't have Chief, but roster does. Looking at this [outdated (May 2018) org chart](https://www.baltimorepolice.org/sites/default/files/General%20Website%20PDFs/BPDOrgChart.pdf), people have "Chief" as a title with their name, but not as an actual job title...I'm collapsing that into the regex as well.

```{r}
# the greatest xpath of my life!
sworn <- read_html("https://en.wikipedia.org/wiki/Baltimore_Police_Department") %>%
  html_node(xpath = "//span[@id='Rank_structure_and_insignia']/parent::h2/following-sibling::table[@class='wikitable']") %>%
  html_table() %>% 
  janitor::clean_names() %>%
  separate_rows(title, sep = "/") %>%
  mutate(title = title %>% str_to_lower() %>% str_trim())

sworn_regex <- c(sworn$title, "chief") %>%
  sort() %>%
  paste(collapse = "|") %>%
  sprintf("(%s)", .)
```

Adding in `is_sworn_officer` code (0 or 1 so it translates easily into Python). 15 jobs are sworn officers.

```{r}
job_code_lookup <- rost_codes %>%
  mutate(is_sworn_officer = str_detect(title_clean, sworn_regex) %>% as.numeric()) %>%
  distinct(job_code, title_clean, is_sworn_officer)

job_code_lookup %>%
  filter(is_sworn_officer == 1)
```

CSV file is in `job_codes.csv`.

```{r}
write_csv(job_code_lookup, "job_codes.csv")
```


